---
title: "Project proposal"
author: "charliebitmyfinger"
output: html_document
---

Accessibility:

alt text.... {r , fig.alt = "ur alt text"}

scale_colour_viridis_d() or scale_fill_viridis_d()

write up: readme description of the project - pointing out most important findings (kinda like https://dcs-210.github.io/project-matt-annie-ethan/)


```{r load-packages, message = FALSE, echo = FALSE}
library(tidyverse)
library(broom)
library(lubridate)
library(scales)
```

## 1. Introduction

We aim to examine what makes a Youtube video popular using a dataset that contains Youtube data.

```{r read-in-data, echo = FALSE, message = FALSE}
yt <- readr::read_csv(file = "../data/USvideos.csv")
```
The data is from Kaggle (the Kaggle link can be found here... (https://www.kaggle.com/datasnaek/youtube-new?select=USvideos.csv) . It was posted to Kaggle by a user called Mitchell J. Mitchell J was able to build the dataset by writing a Python script that scraped the web for Youtube data (scraper code can be found here.... https://github.com/mitchelljy/Trending-YouTube-Scraper) . The Kaggle post includes data on Youtube statistics for USA, Great Britain, Germany, Canada, France, Russia, Mexico, South Korea, Japan, and India (each country's is posted in a separate dataset). We decided to only look at US data for two reasons. 

Reason #1: the datasets all together (if combined) would have been enormous. With so many megabytes of data being processed by our code, our program may have been a bit slower with all countries' data included.
Reason #2: the US is the country we know the best. We may have an easier time drawing conclusions and designing analysis of US data than we would with world data because we are American researchers that can relate to American search habits.

There are `ncol(yt)` columns and `nrow(yt)` rows in the dataset.

Below is a list of the variables and what they mean....
video_id: a unique alphanumeric id for each video (there are only 6351 unique videos in the data)
trending_date: a date on which that video was trending
video_title: title of the video (a string)
channel_title: name of the Youtube channel (a string)
category_id: number corresponding to category of the video (ranging 1-43)
publish_time: a datatime object for when the video was uploaded to Youtube
tags: categorical tag names attached to the video
views: number of views
likes: number of likes
dislikes: number of dislikes
comment_count: number of comments
thumbnail_link: a link to the video
comments_disabled: boolean for whether or not comments are disabled
ratings_disabled: boolean for whether or not likes and dislikes are disabled
video_error_or_removed: boolean for whether or not the video has been taken off the site
description: description of the video

## 2. Data

```{r data-glimpse, echo = FALSE, message = FALSE}
glimpse(yt)
unique(yt$category_id)
head(yt)
```

We noticed that our dataset contains a variable called `category_id` but the variable doesn't tell us what these categories are. Searching online, we found that Youtube has 16 unique categories, and when using the `unique()` function with our `category_id` variable, they match up. The list of categories can be found here: https://techpostplus.com/youtube-video-categories-list-faqs-and-solutions/. It would be interesting to see how popularity changes with category. In order to create some pretty cool visualizations, we will add a new variable to our dataset called `category`, giving us the name of each corresponding category.

```{r new-variable, echo = FALSE, message = FALSE}
yt$category <- 0

film_animation <- "1"
autos_vehicles <- "2"
music <- "10"
pets_animals <- "15"
sports <- "17"
travel_events <- "19"
gaming <- "20"
people_blogs <- "22"
comedy <- "23"
entertainment <- "24"
news_politics <- "25"
howto_style <- "26"
education <- "27"
science_tech <- "28"
nonprofits_activism <- "29"

for(i in 1:nrow(yt)){
  evaluate_category_id <- yt$category_id[i]
  if(evaluate_category_id %in% film_animation){
    yt$category[i] <- "Film & Animation"
  } else if(evaluate_category_id %in% autos_vehicles){
    yt$category[i] <- "Autos & Vehicles"
  } else if(evaluate_category_id %in% music){
    yt$category[i] <- "Music"
  } else if(evaluate_category_id %in% pets_animals){
    yt$category[i] <- "Pets & Animals"
  } else if(evaluate_category_id %in% sports){
    yt$category[i] <- "Sports"
  } else if(evaluate_category_id %in% travel_events){
    yt$category[i] <- "Travel & Events"
  } else if(evaluate_category_id %in% gaming){
    yt$category[i] <- "Gaming"
  } else if(evaluate_category_id %in% people_blogs){
    yt$category[i] <- "People & Blogs"
  } else if(evaluate_category_id %in% comedy){
    yt$category[i] <- "Comedy"
  } else if(evaluate_category_id %in% entertainment){
    yt$category[i] <- "Entertainment"
  } else if(evaluate_category_id %in% news_politics){
    yt$category[i] <- "News & Politics"
  } else if(evaluate_category_id %in% howto_style){
    yt$category[i] <- "Howto & Style"
  } else if(evaluate_category_id %in% education){
    yt$category[i] <- "Education"
  } else if(evaluate_category_id %in% science_tech){
    yt$category[i] <- "Science & Technology"
  } else if(evaluate_category_id %in% nonprofits_activism){
    yt$category[i] <- "Nonprofits & Activism"
  }else {
    yt$category[i] <- "Other"
  }}
head(yt)
```


# 3. Data analysis plan


```{r data-convert, echo = FALSE, message = FALSE}
yt$trending_date <- as.Date(yt$trending_date, "%y.%d.%m")
```

```{r data-date-range, echo = FALSE, message = FALSE}
range(yt$trending_date) #range is wrong bc r does not understand the yy.dd.mm structure
#we need to reformat the dates to be mm.dd.yy
```

```{r range-glimpse, echo = FALSE, message = FALSE}
range(yt$trending_date) # range is wrong bc R does not understand the yy.dd.mm structure
# we need to reformat the dates to be mm.dd.yy
```

There are a number of variables that we may choose as our dependent variable. Number of likes is fascinating to look at. Number of views is fascinating to look at, too.

We can play around with some or all of our predictor variables in order to investigate what factors lead to a popular Youtube video. Videos that have many tags? Videos in a certain category? Videos on a channel that is already popular? Videos posted at a certain time of day? There are many fun questions that we may ask as we progress through the analysis. All of the 14 variables, other than views and likes (which will be used as dependent variables), will probably be useful at some point in our analysis except for `thumbnail_link` (we won't be doing any web scraping ourselves). `Comment_count`, `tags`, `category`, `channel_title`, `publish_time`, and others will certainly be interesting to examine once we dive into the data. Number of dislikes is an especially juicy variable - what if increased dislikes (indicating increased controversy) are tied to an increase in number of views?

We may want to create our own variables at certain points in our analysis using the `mutate()` function. Further, we may want to create a number of plots to visualize our data using ggplot's `geom_point()`, `geom_bar()`, and other functions. We may want to select certain rows to look at using `select()`. We may want to count the occurrence of certain data trends using `count()`. We may want to sort the data using `arrange()`. There are many other functions that we have learned (and new ones that we will learn) that will be helpful for this project.

Since we do not know which predictor variables (or even which dependent variable) we will use yet, it is impossible to surmise what will suffice as conclusive evidence for our research question. Strong correlation plots and other visualizations that display clear trends will most likely be our "proof" once we come up with some conclusions to include in our final file.

Some visualizations of the data are shown below....

## Views vs. Likes

```{r views-v-likes, echo = FALSE, message = FALSE}
#views v likes
ggplot(data = yt, mapping = aes(x = likes, y = views, color = category)) +
  geom_point(size = 0.5, alpha = 0.5) + 
  labs( "Views vs. Likes", x="Likes", y="Views", color = "Category") + 
  scale_y_continuous(labels = label_number(suffix = " M", scale = 1e-6)) +
  scale_x_continuous(labels = label_number(suffix = " M", scale = 1e-6)) +
  scale_colour_viridis_d()


#views v likes all w line of fit
ggplot(data = yt, mapping = aes(x = likes, y = views, color = category)) +
  geom_point(size = 0.5, alpha = 0.5) + 
  labs( "Views vs. Likes", x="Likes", y="Views", color = "Category") + 
  geom_smooth(method = "lm", se = FALSE, color = "red") + 
  scale_y_continuous(labels = label_number(suffix = " M", scale = 1e-6)) +
  scale_x_continuous(labels = label_number(suffix = " M", scale = 1e-6)) +
  scale_colour_viridis_d()

#views v likes faceted w lines of fit
ggplot(data = yt, mapping = aes(x = likes, y = views, color = category), alpha = 0.1) +
  geom_point(alpha = 0.5) + 
  labs( "Views vs. Likes", x="Likes", y="Views", color = "Category") + 
  scale_y_continuous(labels = label_number(suffix = " M", scale = 1e-6)) +
  geom_smooth(method = "lm", se = FALSE, color = "red") + 
  facet_wrap(~category) +
  scale_x_continuous(labels = label_number(suffix = " M", scale = 1e-6)) +
  scale_colour_viridis_d()
```






## Views vs. Dislikes

```{r views-v-dislikes, echo = FALSE, message = FALSE,}
#views v dislikes
ggplot(data = yt, mapping = aes(x = dislikes, y = views, color = category)) +
  geom_point(size = 0.5, alpha = 0.5) + 
  labs( "Views vs. Dislikes", x="Dislikes", y="Views", color = "Category") + 
  scale_y_continuous(labels = label_number(suffix = " M", scale = 1e-6)) +
  scale_x_continuous(labels = label_number(suffix = " M", scale = 1e-6)) +
  scale_colour_viridis_d()

#views v dislikes all w line of fit
ggplot(data = yt, mapping = aes(x = dislikes, y = views, color = category)) +
  geom_point(size = 0.5, alpha = 0.5) + 
  labs( "Views vs. Dislikes", x="Dislikes", y="Views", color = "Category") + 
  geom_smooth(method = "lm", se = FALSE, color = "red") + 
  scale_y_continuous(labels = label_number(suffix = " M", scale = 1e-6)) +
  scale_x_continuous(labels = label_number(suffix = " M", scale = 1e-6)) +
  scale_colour_viridis_d()

#views v dislikes faceted w lines of fit
ggplot(data = yt, mapping = aes(x = dislikes, y = views, color = category), alpha = 0.1) +
  geom_point(alpha = 0.5) + 
  labs( "Views vs. Dislikes", x="Dislikes", y="Views", color = "Category") + 
  scale_y_continuous(labels = label_number(suffix = " M", scale = 1e-6)) +
  geom_smooth(method = "lm", se = FALSE, color = "red") + 
  facet_wrap(~category) +
  scale_x_continuous(labels = label_number(suffix = " M", scale = 1e-6)) +
  scale_colour_viridis_d()
```

## Views vs. Comment Count

```{r views-v-comments, echo = FALSE, message = FALSE}
#views v comments
ggplot(data = yt, mapping = aes(x = comment_count, y = views, color = category)) +
  geom_point(size = 0.5, alpha = 0.5) + 
  labs( "Views vs. Comments", x="Comment Count", y="Views", color = "Category") + 
  scale_y_continuous(labels = label_number(suffix = " M", scale = 1e-6)) +
  scale_x_continuous(labels = label_number(suffix = " M", scale = 1e-6)) +
  scale_colour_viridis_d()


#views v comments all w line of fit
ggplot(data = yt, mapping = aes(x = comment_count, y = views, color = category)) +
  geom_point(size = 0.5, alpha = 0.5) + 
  labs( "Views vs. Comments", x="Comment Count", y="Views", color = "Category") + 
  geom_smooth(method = "lm", se = FALSE, color = "red") + 
  scale_y_continuous(labels = label_number(suffix = " M", scale = 1e-6)) +
  scale_x_continuous(labels = label_number(suffix = " M", scale = 1e-6)) +
  scale_colour_viridis_d()

#views v comments faceted w lines of fit
ggplot(data = yt, mapping = aes(x = comment_count, y = views, color = category), alpha = 0.1) +
  geom_point(alpha = 0.5) + 
  labs( "Views vs. Comments", x="Comment Count", y="Views", color = "Category") + 
  scale_y_continuous(labels = label_number(suffix = " M", scale = 1e-6)) +
  geom_smooth(method = "lm", se = FALSE, color = "red") + 
  facet_wrap(~category) +
  scale_x_continuous(labels = label_number(suffix = " M", scale = 1e-6)) +
  scale_colour_viridis_d()
```

## Views vs. Date Published

```{r views-v-date-published, echo = FALSE, message = FALSE, warning = FALSE}
#views v date published
yt$publish_date <- as.Date(yt$publish_time)

yt %>%
  ggplot(data = yt, mapping = aes(x = publish_date, y = views, color = category)) +
  geom_point(size = 0.5, alpha = 0.5) +
  scale_y_continuous(labels = label_number(suffix = " M", scale = 1e-6)) +
  labs(
    title = "The Relationship Between YouTube Views and Date",
    subtitle = "by Category",
    x = "Date Published",
    y = "Views",
    color = "Category") +
  theme(legend.title = element_text(size = 10), 
        legend.text = element_text(size = 8),
        legend.key.size = unit(0.4, "cm"), 
        axis.text.x=element_text(angle=60, hjust=1)) +
  scale_x_date(limit=c(as.Date("2016-12-15 18:59:30 UTC"), as.Date("2018-6-30 18:59:30 UTC")), date_breaks = "1 month", date_labels = "%y-%m") +
  scale_colour_viridis_d()

#views v dates all w line of fit
yt %>%
  ggplot(data = yt, mapping = aes(x = publish_date, y = views, color = category)) +
  geom_point(size = 0.5, alpha = 0.5) +
  scale_y_continuous(labels = label_number(suffix = " M", scale = 1e-6)) +
  labs(
    title = "The Relationship Between YouTube Views and Date",
    subtitle = "by Category",
    x = "Date Published",
    y = "Views",
    color = "Category") +
  theme(legend.title = element_text(size = 10), 
        legend.text = element_text(size = 8),
        legend.key.size = unit(0.4, "cm"), 
        axis.text.x=element_text(angle=60, hjust=1)) +
  scale_x_date(limit=c(as.Date("2016-12-15 18:59:30 UTC"), as.Date("2018-6-30 18:59:30 UTC")), date_breaks = "1 month", date_labels = "%y-%m") +
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  scale_colour_viridis_d()

#views v dates faceted w lines of fit
yt %>%
  ggplot(data = yt, mapping = aes(x = publish_date, y = views, color = category)) +
  geom_point(alpha = 0.5) +
  labs(
    title = "The Relationship Between YouTube Views and Date",
    subtitle = "by Category",
    x = "Date Published",
    y = "Views",
    color = "Category") +
  theme(legend.title = element_text(size = 10), 
        legend.text = element_text(size = 8),
        legend.key.size = unit(0.4, "cm"), 
        axis.text.x=element_text(angle=60, hjust=1)) +
  scale_x_date(limit=c(as.Date("2017-10-15 18:59:30 UTC"), as.Date("2018-6-30 18:59:30 UTC")), date_breaks = "1 month", date_labels = "%y-%m") +
  scale_y_continuous(labels = label_number(suffix = " M", scale = 1e-6)) +
  facet_wrap(~category) +
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  scale_colour_viridis_d()
```


```{r mr-wan-time-plots, echo = FALSE, message = FALSE, warning = FALSE}
ggplot(data = yt, mapping = aes(x = month(publish_time, label = TRUE), y = views)) +
  geom_point(size = 0.5) + 
  coord_polar() + 
  labs(title = "Views and Upload Time (Month)", x="Month", y="Views") + 
  scale_y_continuous(labels = label_number(suffix = " M", scale = 1e-6))
                                                    
                             
ggplot(data = yt, mapping = aes(x = hour(publish_time), y = views)) + 
  geom_point(size = 0.5) + 
  coord_polar() + 
  labs(title = "Views and Upload Time(Hour)", x="Hour", y="Views") + 
  scale_y_continuous(labels = label_number(suffix = " M", scale = 1e-6))
```

## Views v Number Of Characters In Video Title

```{r length-of-title, echo = FALSE, message = FALSE, warning = FALSE}
yt = yt %>%
  mutate(title_length = nchar(title)) #the number of characters in the title


#length of title v views
ggplot(data = yt, mapping = aes(x = title_length, y = views, color = category)) +
  geom_point(size = 0.5, alpha = 0.5) +
 theme(legend.title = element_text(size = 10), 
        legend.text = element_text(size = 8),
        legend.key.size = unit(0.4, "cm"), 
        axis.text.x=element_text(angle=60, hjust=1)) +
  scale_colour_viridis_d() +
  labs(
    title = "The Relationship Between YouTube Views and # of Charcters in the Video Title",
    subtitle = "by Category",
    x = "# of Characters in the Video Title",
    y = "Views",
    color = "Category") + scale_y_continuous(labels = label_number(suffix = " M", scale = 1e-6))

#length of title v views with line of best fit
ggplot(data = yt, mapping = aes(x = title_length, y = views, color = category)) +
  geom_point(size = 0.5, alpha = 0.5) +
 theme(legend.title = element_text(size = 10), 
        legend.text = element_text(size = 8),
        legend.key.size = unit(0.4, "cm"), 
        axis.text.x=element_text(angle=60, hjust=1)) +
  scale_colour_viridis_d() +
  labs(
    title = "The Relationship Between YouTube Views and # of Charcters in the Video Title",
    subtitle = "by Category",
    x = "# of Characters in the Video Title",
    y = "Views",
    color = "Category") + scale_y_continuous(labels = label_number(suffix = " M", scale = 1e-6))+
  geom_smooth(method = "lm", se = FALSE, color = "red")



#length of title v views with line of best fit faceted
ggplot(data = yt, mapping = aes(x = title_length, y = views, color = category)) +
  geom_point() +
  facet_wrap(~category) +
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  theme(legend.title = element_text(size = 10), 
        legend.text = element_text(size = 8),
        legend.key.size = unit(0.4, "cm"), 
        axis.text.x=element_text(angle=60, hjust=1), panel.spacing = unit(0.1, "lines"),
      axis.ticks.x=element_blank()) +
  scale_colour_viridis_d() +
  labs(
    title = "The Relationship Between YouTube Views And # Of Charcters In The Video Title",
    subtitle = "by Category",
    x = "# of Characters in the Video Title",
    y = "Views",
    color = "Category") + scale_y_continuous(labels = label_number(suffix = " M", scale = 1e-6))+
  geom_smooth(method = "lm", se = FALSE, color = "red")
```

## Category Analysis

```{r videos-per-category, echo = FALSE, message = FALSE}
videos_per_category = yt %>%
    group_by(category) %>%
    count %>%
    arrange(desc(n))

ggplot(videos_per_category, aes(x=reorder(category, n), y=n)) + 
  geom_bar(stat = "identity") +
  labs(title = "Number of Videos In Each Category", x="Category", y="# of Videos") +
   theme(axis.text.x=element_text(angle=60, hjust=1), legend.position= 'none')
```

```{r cateogry-v-views, echo = FALSE, message = FALSE}
yt %>%
  ggplot(data = yt, mapping = aes(x = category, y = views)) +
  geom_boxplot(alpha = 0.15, aes(color = factor(yt$category))) +
  labs(
    title = "Relationship Between YouTube Category and Views",
    x = "Category",
    y = "Views",
    color = "Category") +
  theme(axis.text.x=element_text(angle=60, hjust=1), legend.position= 'none' ) +
  scale_colour_viridis_d() + 
  ylim(0, 235000000) + scale_y_continuous(labels = label_number(suffix = " M", scale = 1e-6))
```

```{r cateogry-v-views-mean, echo = FALSE, message = FALSE}

views_for_category = yt %>%
  group_by(category) %>%
  summarise(mean_views = mean(views), median_views = median(views)) %>%
  arrange(desc(mean_views))

ggplot(data = views_for_category) +
  geom_segment(aes(x=reorder(category, mean_views), xend=reorder(category, mean_views), y=median_views, yend=mean_views)) + 
  geom_point( aes(x=reorder(category, mean_views), y=median_views), color= "dark green", size=2 ) +
  geom_point( aes(x=reorder(category, mean_views), y=mean_views), color= "purple", size=2 ) +
 theme(legend.position = 'none') +
  coord_flip() + 
  scale_y_continuous(labels = label_number(suffix = " M", scale = 1e-6)) +
  labs( title = "Mean and Median Views For Each Category", subtitle = "green is median, purple is mean", y=" Number of Views", x="Category")

```

## Other Plots

```{r max-views-by-channel, echo = FALSE, message = FALSE}
max_views_vid_for_each_channel = yt %>%
  group_by(channel_title) %>%
  summarise(max_views = max(views)) %>%
  arrange(desc(max_views))


#in honor of dan mills....
ggplot(data = max_views_vid_for_each_channel, mapping = aes(x = reorder(channel_title, max_views), y = max_views)) +
  geom_point(size = 0.75) + 
  theme(axis.text.x=element_text(angle=90, hjust=1), legend.position= 'none' ) +
  scale_y_continuous(labels = label_number(suffix = " M", scale = 1e-6)) +
  labs( "Max Views Video For Each Channel", x="Channel Title", y="# of Views on Most-Viewed Video")
```

```{r trending-gap, echo = FALSE, message = FALSE, warning = FALSE}
#the trending gap is the amount of days between when a video was published and when it was trending

yt = yt %>%
  mutate(trending_gap = as.numeric(trending_date - publish_date))

ggplot(data = yt, mapping = aes(x = trending_gap, y = views, color = category)) +
  geom_point(alpha = .7) +
 theme(legend.title = element_text(size = 10), 
        legend.text = element_text(size = 8),
        legend.key.size = unit(0.4, "cm"), 
        axis.text.x=element_text(angle=60, hjust=1)) +
  scale_colour_viridis_d() +
  labs(
    title = "The Relationship Between YouTube Views And # Of Days Between The Date A Video Was Published And The Date It Was Trending",
    subtitle = "by Category",
    x = "# Of Days Between The Date A Video Was Published And The Date It Was Trending",
    y = "Views",
    color = "Category") + scale_y_continuous(labels = label_number(suffix = " M", scale = 1e-6))
```

The takeaway from the plot above is that the videos that do garner many views tend to begin trending shortly after they are published.

```{r videos-per-channel, echo = FALSE, message = FALSE, warning = FALSE}
#do videos from channels that have more videos tend to get more views?

videos_per_channel = yt %>%
    group_by(channel_title) %>%
    count %>%
    arrange(desc(n))

mean_views_vid_for_each_channel = yt %>%
  group_by(channel_title) %>%
  summarise(mean_views = mean(views)) %>%
  arrange(desc(mean_views))

ggplot(mapping = aes(x = videos_per_channel$n, y = mean_views_vid_for_each_channel$mean_views)) +
  geom_point(size = 0.75) +
  theme(axis.text.x=element_text(angle=90, hjust=1), legend.position= 'none') +
  labs(title = "Mean Views Per Channel In Comparison to Number of Videos Per Channel",
    x = "The Number of Videos Per Channel",
    y = "The Mean Number of Views Per Video for Each Channel")
```